---
title: "Appendices for 'An experimental test of the Allee effect range limitation hypothesis'"
author: Samuel Merker and Richard Chandler
fig_width: 6 
fig_height: 6 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Appendix 1: Global Model

JAGS code for the dynamic N-mixture model including an interaction of climate and treatment effects on Canada warbler density and growth rate in the southern Appalachian Mountains of the United States.

``` {r jags-model,eval=FALSE}
model {
    
    ## Coefficients of model for initial abundance
    beta0 ~ dnorm(0, 1)    
	beta1 ~ dnorm(0, 0.1)
    ## Coefficients of model for detection probability
	alpha0 ~ dnorm(0, 0.5) 
	alpha1 ~ dnorm(0, 0.1)
	alpha2 ~ dnorm(0, 0.1)
	alpha3 ~ dnorm(0, 0.1)
	alpha4 ~ dnorm(0, 0.1)
    ## Coefficients of model for growth rate
	lambda0 ~ dnorm(0, 1)   
	lambda1 ~ dnorm(0, 0.1)
	lambda2 ~ dnorm(0, 0.1)
	lambda3 ~ dnorm(0, 0.1)
	lambda4 ~ dnorm(0, 0.1)
	lambda5 ~ dnorm(0, 0.1)
	lambda6 ~ dnorm(0, 0.1)
	for(i in 1:nSites) {
      ## Model for year 1
	  log(psi[i,1]) <- beta0 + beta1*climate[i]  ## psi is the expected value of N
	  D[i,1] <- psi[i,1] / plotArea              ## D is the expected value of density
	  DS[i,1] <- (D[i,1]-0.15)/0.25              ## Density (roughly) standardized
	  N[i,1] ~ dpois(psi[i,1])                   ## Realized value of abundance
	  ld.N.site.year[i,1] <- logdensity.pois(N[i,1], psi[i,1]) ## for WAIC
	  for(j in 1:4) {
        ## Detection Probability  
        logit(p[i,j,1]) <- alpha0 + alpha1*wind[i,j,1] + alpha2*noise[i,j,1] +
              alpha3*date[i,j,1] + alpha4*time[i,j,1] 
        ## Count Data
        y[i,j,1] ~ dbin(p[i,j,1], N[i,1])
        ## Residuals  
        resid[i,j,1] <- y[i,j,1]-N[i,1]*p[i,j,1]
        ## Log-densities to be used for WAIC  
	    ld.y.site.year.visit[i,j,1] <- logdensity.bin(y[i,j,1], p[i,j,1], N[i,1])
	    ld.yN.site.year.visit[i,j,1] <- ld.y.site.year.visit[i,j,1]+ld.N.site.year[i,1]
	  }
	  ld.yN.site.year[i,1] <- ld.N.site.year[i,1]+sum(ld.y.site.year.visit[i,,1])
      resid.site.yr[i,1] <- mean(resid[i,1:4,1]) ## Average residuals at each site
        
      ## Model for years 2, 3, 4
      for(t in 2:T) {
        ## Growth Rate  
        log(lambda[i,t-1]) <- lambda0 + lambda1*climate[i] + lambda2*treatment[i,1,t] +
            lambda3*DS[i,t-1] + lambda4*DS[i,t-1]^2 + lambda5*DS[i,t-1]*climate[i] +
            lambda6*DS[i,t-1]^2*climate[i] 
        psi[i,t] <- psi[i,t-1]*lambda[i,t-1]  
        D[i,t] <- psi[i,t] / plotArea
        DS[i,t] <- (D[i,t]-0.15)/0.25              
	    N[i,t] ~ dpois(psi[i,t])
	    ld.N.site.year[i,t] <- logdensity.pois(N[i,t], psi[i,t])
        for(j in 1:nIntervals[t]) {
          logit(p[i,j,t]) <- alpha0 + alpha1*wind[i,j,t] + alpha2*noise[i,j,t] +
                alpha3*date[i,j,t] + alpha4*time[i,j,t]
          y[i,j,t] ~ dbin(p[i,j,t], N[i,t])
          resid[i,j,t] <- y[i,j,t]-N[i,t]*p[i,j,t]
          ld.y.site.year.visit[i,j,t] <- logdensity.bin(y[i,j,t], p[i,j,t], N[i,t])
          ld.yN.site.year.visit[i,j,t] <- ld.y.site.year.visit[i,j,t]+ld.N.site.year[i,t]
	    }
        ld.yN.site.year[i,t] <- ld.N.site.year[i,t]+sum(ld.y.site.year.visit[i,,t])
        resid.site.yr[i,t] <- mean(resid[i,1:4,t])
	  }
	  ld.yN.site[i] <- sum(ld.yN.site.year[i,])
	}
	for(t in 1:T) {
	  Ntotal[t] <- sum(N[,t])
	}
}
```


\newpage

## Appendix 2

An assessment of spatial autocorrelation in the residuals using Moran's I.

### Moran's I
```{r moransI}
library(spdep)
cawa.resids <- read.csv("cawa-pb-resid.csv", row.names=1)
load("points.R") 
## Coordinates (in the same site-order as resids)
coords <- as.matrix(all.points.xy)
dist.mat <- as.matrix(dist(coords))    # Distance matrix
weights <- ifelse(dist.mat>1000, 0, 1) # Neighbors defined as sites within 1 km
diag(weights) <- 0                     # Weight=1 for neighbors, 0 otherwise
weight.list <- mat2listw(weights)      # create neighbor list for spdep
Mi.1 <- moran.test(cawa.resids[,1], weight.list) #test of year 1 resids
Mi.2 <- moran.test(cawa.resids[,2], weight.list) #test of year 2 resids
Mi.3 <- moran.test(cawa.resids[,3], weight.list) #test of year 3 resids
Mi.4 <- moran.test(cawa.resids[,4], weight.list) #test of year 4 resids
```

Table 1. Moran's I for each year shows no indication of spatial autocorrelation in the model residuals.
```{r kable, echo=FALSE, results='asis'}
library(knitr)
MoranI.tbl <- data.frame(
    Year=c(1,2,3,4), 
    "Moran's I"=c(Mi.1$statistic[1], Mi.2$statistic[1],Mi.3$statistic[1], Mi.4$statistic[1]), 
    "p-value"=c(Mi.1$p.value[1], Mi.2$p.value[1],Mi.3$p.value[1], Mi.4$p.value[1]))
kable(MoranI.tbl, digits=3, col.names=c("Year", "Moran's I", "p-value"))
```


```{r resid-trans, include=FALSE}
cawa.resids <- read.csv("cawa-pb-resid.csv")
load("points.R")
# resids.pos <- cawa.resids[,2:5]+(0.6493417)*2.5
resids.pos <- (cawa.resids[,2:5]+0.65)*1.5
min(resids.pos)
```
\newpage

## Appendix 3

Model residuals at each survey point. The size of the point is proportional to the residual. 

```{r resid-maps,echo=FALSE,fig.height = 6,fig.width = 8,fig.show="hold",out.width="50%"}
plot(all.points.xy, type="p",pch=16, cex=resids.pos[,1], asp=1,col="black",
     xlab="", xlim=c(274500, 280000), main="2014")
plot(all.points.xy, type="p",pch=16, cex=resids.pos[,2], asp=1,col="black",
     xlab="", xlim=c(274500, 280000), main="2015")
plot(all.points.xy, type="p",pch=16, cex=resids.pos[,3], asp=1,col="black",
     xlab="", xlim=c(274500, 280000), main="2016")
plot(all.points.xy, type="p",pch=16, cex=resids.pos[,4], asp=1,col="black",
     xlab="", xlim=c(274500, 280000), main="2017")
```
\newpage

## Appendix 4

Breeding range of Canada Warbler.
```{r make-cawa-map, echo=FALSE, results="hide", warning=FALSE}
###CAWA map for Allee-effects paper###
library(raster)
library(rgdal)
library(rworldmap)
library(maptools)
library(maps)
library(cleangeo)
library(rgeos)
library(GISTools)
sPDF<-getMap()
sPDF <- clgeo_Clean(sPDF)  ## Needed to fix up some non-closed polygons 
cont <-sapply(levels(sPDF$continent),
              FUN = function(i) {
                ## Merge polygons within a continent
                poly <- gUnionCascaded(subset(sPDF, continent==i))
                ## Give each polygon a unique ID
                poly <- spChFIDs(poly, i)
                ## Make SPDF from SpatialPolygons object
                SpatialPolygonsDataFrame(poly,
                                         data.frame(continent=i, row.names=i))
              },
              USE.NAMES=TRUE)
## Bind the 6 continent-level SPDFs into a single SPDF
cont <- Reduce(spRbind, cont)
cont
N.A<-subset(cont, continent=="North America")
CAWA.map<-readOGR("C:/Users/ChandlerLab/Desktop/Dissertation/Egg Swapping/Egg Swapping Chapter/maps/Cardellina_canadensis_2_4434.shp")
#CAWA.map<-readOGR("Cardellina_canadensis_2_4434.shp")
```


```{r cawa-map, echo=FALSE, fig.height = 6, fig.width = 8, fig.align = "center"}
par(oma=c(1,1,1,1))
plot(N.A, xlim=c(-100,-90), ylim=c(26.5,68))
plot(CAWA.map, col="green", add=TRUE)
#points(-84.2,35, pch=22, bg="white",col="black")
arrows(-93,37, -84.7,35.5, length=0.1, lwd=2)
north.arrow(-130,38,len=1.2,lab='NORTH', lwd=0.7, cex.lab=0.7,tcol='black')
text(-105,38, c("Traling-edge of breeding range \n Nantahala National Forest \n North Carolina, USA"), bty='n', cex=.6)
box(which = "plot", lty = "solid")
raster::scalebar(d = 1000, # distance in km
                 xy = c(-70, 35),
                 type = "bar", 
                 divs = 2, 
                 below = "km", 
                 lonlat = TRUE,
                 label = c(0,50,100), 
                 # adj=c(0, -0.75), 
                 lwd = 2)
```

